\section{Man against Machine}
%\chapter{Man against Machine}
\label{sect:mam_study}
Here we set an experienced human scheduler up against a series of automated schedulers to see who wins.
description - likely problems assumptions errors how handled, input forms, what they did - notes from subject, how measured, results-comparison with other schedulers.

\subsection{Rational and introduction}
Seelcting appropriate groups to perform is a complex task. There are many tardeoffs to be made between the various competing preferences and constraints. When the schedule is heavily loaded, i.e. there are potentially more observations to perform than could be accomodated even under ideal conditions, the job of the scheduler becomes extremely difficult. A major problem is to determine just what those tradeoffs are - i.e. just how much more important is a high priority, non-urgent group relative to a medium priority urgent group?. It is because it is difficult to quantify these relative weights that we turn to the human scheduler. A human makes these sort of trade-offs on a daily basis - often with little knowledge they are doing it - if we set a human up to perform the task we might be able to deduce from what they have done and the choices they have made some of the rules they are using (often without realizing) and the relative weighting they employ. Additionally it was thought that it would be useful just to know if a human was in fact capable of \emph{beating} an automated scheduler in terms of the standard SQMs.

\subsection{Method}

\clearpage

\begin{landscape}

\begin{table}
\begin{center}
\caption{Example output from HumanSchedulerTestGenerator}
\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|}
\hline
{\bf ID} & {\bf RA}    & {\bf Dec} & {\bf Name} & {\bf Timing} & {\bf Moon}  & {\bf Seeing} & {\bf Night}   & {\bf Priority}    & {\bf Exec}   & {\bf Urgency} \\
\hline
G\_448  &22:41:55.90  &20:15:42.00  &UCM\_galaxies\_6&FLEXBL  &DARK &AVER  &       &1 &39.63M  &11 \\
\hline
G\_449  &23:51:25.70  &24:24:12.00  &UCM\_galaxies\_18&FLEXBL &DARK &AVER  &       &1 &39.63M  &12 \\
\hline
G\_450  &23:22:20.90  &23:0:42.00&UCM\_galaxies\_14  &FLEXBL  &DARK &AVER  &       &1 &39.63M  &12 \\
\hline
G\_452  &22:58:50.00  &20:17:54.00  &UCM\_galaxies\_7&FLEXBL  &DARK &AVER  &       &1 &39.63M  &11 \\
\hline
G\_453  &23:31:48.50  &24:44:6.00&UCM\_galaxies\_17  &FLEXBL  &DARK &AVER  &       &1 &39.63M  &12 \\
\hline
G\_455  &23:5:25.50&21:9:42.00&UCM\_galaxies\_4&FLEXBL        &DARK &AVER  &       &1 &39.63M  &12  \\
\hline
G\_466  &22:53:57.74  &16:8:53.56&2251  &MONITR 168H [167]    &DARK &AVER  &       &1 &7.8M   &2  \\
\hline
G\_469  &22:2:43.291  &42:16:39.98  &2200 &MONITR 168H [167]  &     &AVER  &       &1 &7.8M   &2 \\ 
\hline
G\_470  &2:38:38.93&16:36:59.28  &0235 &MONITR 168H [167]     &DARK &AVER  &       &1 &7.8M   &2  \\
\hline
G\_471  &12:29:6.70&2:3:8.60  &1226 &MONITR 168H [167]        &     &AVER  &       &1 &7.8M   &2  \\
\hline
G\_472  &8:54:48.90&20:6:30.64&0851 &MONITR 168H [167]        &DARK &AVER  &       &1 &7.8M   &2  \\
\hline
G\_476  &0:42:49.64&41:15:26.50  &ANGM31  &MONITR 2.5H [2]    &     &POOR  &       &2 &33.4M   &CRIT  \\
\hline
G\_481  &14:17:59.60  &25:6:53.00&5548\_opt&MONITR 168H [120] &     &POOR  &ASTR   &2 &3.13M  &3  \\
\hline
\end{tabular}
\end{center}
\end{table}

\end{landscape}



\subsection{Results}

\subsection{Conclusions}
